Revisiting the hydrodynamic modulation of short waves by longer waves

### 2024-09-06 paper/paper.tex
Analytical solutions done

\begin{equation}
\label{eq:k_short_exact}
\widetilde{k} = k e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
\end{equation}

@@

\begin{equation}
\label{eq:k_short_lhs}
\widetilde{k} = k (1 + \varepsilon_L \cos{\psi})
\end{equation}
Coming from the conservation of wave crests, as opposed to the velocity potential
superposition approach, the L-HS solution requires two approximations in their
solution:
First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
second, further truncating the Taylor expansion series beyond the first order term.
These approximations cause an underestimate of the short-wave modulation
magnitude (Fig. \ref{fig:analytical_solutions_ak0.1}a).
At the wave crest, Eq. \ref{eq:k_short_exact} yields the wavenumber modulation
that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of \citet{longuet1960changes}
for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
Although not immediately obvious from the expressions, the modulated wavenumber
in Eq. \ref{eq:k_short_lhs} is conservative across the long-wave phase, whereas
the higher-order solution in Eq. \ref{eq:k_short_exact} is not.
This is due to the fact that combining the orbital velocities evaluated at
$z=\eta$ and the linearized conservation of wave crests in Eq. \ref{eq:wavenumber}
results in excess short-wave wavenumber being created.
The solution of the full Eq. \ref{eq:wavenumber} of course must be conservative
across the long-wave phase, as we will see later from the numerical simulations.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.1.png}
\caption{
  Modulation of short-wave (a) wavenumber, (b) gravitational acceleration,
  (c) amplitude, (d) steepness, (e) intrinsic frequency, and (d) intrinsic phase
  speed as function of long-wave phase for $\varepsilon_L = 0.1$, based on
  analytical solutions by \citet{longuet1960changes} (L-HS 1960, black)
  and this paper (blue).
  Long-wave crest and trough are located at $\psi = 0$ and $\psi = \pi$,
  respectively.
}
\label{fig:analytical_solutions_ak0.1}
\end{figure}

@@

\begin{equation}
\label{eq:gravity_modulation_general}
\widetilde{g}
  = g \cos{\alpha} 
  + \dfrac{\partial W_{z=\eta}}{\partial t} \cos{\alpha}
  + \dfrac{\partial U_{z=\eta}}{\partial t} \sin{\alpha}
\end{equation}

@@

\begin{equation}
\label{eq:gravity_modulation_linear}
\widetilde{g}
  = g 
  \frac{
    1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}
    \left[ 1 + \left(\varepsilon_L \sin{\psi}\right)^2 \right]
  }
  {\sqrt{\left(\varepsilon_L \sin{\psi}\right)^2 + 1}}
\end{equation}

@@

\begin{equation}
\label{eq:gravity_modulation_linear_no_curvature}
\widetilde{g} = g \left( 1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}} \right)
\end{equation}
Further, if we evaluate the orbital accelerations at the mean water level $z=0$,
this further simplifies to:

\begin{equation}
\label{eq:gravity_modulation_linear_no_curvature_mean_level}
\widetilde{g} = g \left( 1 - \varepsilon_L \cos{\psi} \right)
\end{equation}
which is the form used by \citet{peureux2021unsteady}, for example.

@@

\begin{equation}
\label{eq:gravity_modulation_stokes}
\widetilde{g}(\psi) =
g \left\{
  1 - \varepsilon_L e^{k \eta_{St}}
  \left[ \cos{\psi} -
    \varepsilon_L \sin{\psi} \left(
      \sin{\psi}
      + \varepsilon_L \sin{2\psi}
      - \dfrac{1}{16} \varepsilon_L^2 \sin{\psi}
      + \dfrac{9}{8} \varepsilon_L^2 \sin{3\psi}
    \right)
  \right]
\right\}
\end{equation}
The equivalent expression that includes the curvilinear effects can be derived
by evaluating orbital accelerations and local slope $\alpha$ at $z = \eta_{St}$,
and inserting them into Eq. \ref{eq:gravity_modulation_general}:

\begin{equation}
\label{eq:gravity_modulation_general_stokes}
\widetilde{g}
  = g \cos{\alpha} 
  + \dfrac{\partial W_{z=\eta_{St}}}{\partial t} \cos{\alpha_{St}}
  + \dfrac{\partial U_{z=\eta_{St}}}{\partial t} \sin{\alpha_{St}}
\end{equation}

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_effective_gravities.png}
\caption{
  (a) Analytical solutions for the effective gravitational acceleration
  modulation by long waves, as function of long-wave phase, for $\varepsilon_L = 0.2$.
  Black is for linear wave evaluated at $z=0$,
  blue and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and without curvature effects,
  and orange and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and with curvature effects.
  (b) As in (a) but normalized by the modulation of the linear wave evaluated at $z=0$.
}
\label{fig:effective_gravities}
\end{figure}

It is instructive to also examine the mean, maximum, and minimum values of the
effective gravity as function of long-wave steepness $\varepsilon_L$
(Fig. \ref{fig:effective_gravities_mean_max_min}).
The long-wave averaged effective gravity of short waves remains equal to the
resting value of $g$ for all values of $\varepsilon_L$, except if the effects
of curvature are considered, in which case the average effective gravity is
reduced.
This reduction, in case of a long Stokes wave, is $\approx 3\%$ for
$\varepsilon_L = 0.2$ and $\approx 14\%$ for $\varepsilon_L = 0.4$.
As the effective gravity minima and maxima occur at crests and troughs,
respectively, the curvature effects only affect the phase-averaged values but
not the extremes.
The effective gravity at the crest of the steepest ($\varepsilon_L = 0.4$) long
Stokes wave is reduced by $\approx 64\%$.

@@

The modulation of short-wave amplitude can be derived in the same way by
dropping the non-linear terms in the wave action balance equation
(Eq. \ref{eq:wave_action2}) and integrating wave action in time:

\begin{equation}
\label{eq:wave_action_modulation}
\widetilde{N}(\psi) = N e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
\end{equation}

Since wave action is energy divided by the intrinsic frequency $\sigma$,
and energy scales with $ga^2$, the modulation of the amplitude is readily
related to the modulations of gravity, wavenumber, and action:

\begin{equation}
\label{eq:wave_amplitude_modulation}
\dfrac{\widetilde{a}}{a} = \sqrt{
  \dfrac{g}{\widetilde{g}}
  \dfrac{\widetilde{\sigma}}{\sigma}
  \dfrac{\widetilde{N}}{N}}
=
  \left( \dfrac{g}{\widetilde{g}} \right)^{0.25}
  \left( \dfrac{\widetilde{k}}{k} \right)^{0.25}
  \left( \dfrac{\widetilde{N}}{N} \right)^{0.5}
\end{equation}
To the 1$^{st}$ order in $\varepsilon_L$, and using Eqs. \ref{eq:k_short_exact}
and \ref{eq:wave_action_modulation}, we get:

@@

@@

We now proceed to integrate the full numerical solutions for the short-wave
wavenumber and action density in presence of long waves, and in absence of
sources or sinks of wave energy.
Equations (\ref{eq:wavenumber})-(\ref{eq:wave_action2}) are discretized using
2$^{nd}$  order central finite difference in space and integrated in time using the
4$^{th}$ order Runge-Kutta method \citep{butcher1996runge}.
The space is divided into 100 grid points.
This is effectively the same equation set and numerical configuration as that of
\citet{peureux2021unsteady}, except that, like \citet{longuet1960changes}, their
long-wave orbital velocities are evaluated at $z=0$ instead of $z=\eta_L$.
The key difference is that evaluating orbital velocities at $z=0$ neglects the
Stokes drift of short-wave groups.
Although not significantly consequential for the modulation results here, there
is no good reason to exclude it given that its implementation is trivial.

Another difference is the choice of the numerical scheme for spatial differences,
which in their case is the MUSCL4 scheme by Kurganov and Tadmor (2000 TODO).
Although the 2$^{nd}$ order central finite difference is not appropriate for many
numerical problems, in our case it is sufficient because it is conservative and
the fields that we compute derivatives of are smooth and continuous.
We show in the next section that our numerical results in infinite long-wave
trains are qualitatively equivalent to those of \citet{peureux2021unsteady}.

Finally, the numerical equations here are integrated in a fixed reference frame,
rather than one that moves with the long wave phase speed.
Either approach produces qualitatively equivalent results, but we find that
a fixed reference frame allows for more intuitive interpretation of the results.
However, this difference is important to keep in mind when comparing the results
of \citet{peureux2021unsteady} and those presented here--in their Fig. 1 the
long-wave phase is fixed and the short waves are moving leftward
with the speed of $C_{pL} - C_g - U$ (neglecting any group speed inhomogeneity),
whereas in the figures that follow, the long wave is moving rightward with its
phase speed $C_{pL}$ and the short waves (that is, their energy) are moving
rightward with the speed of $C_g + U$ (neglecting any group wave inhomogeneity).

@@

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{figures/numerical_vs_analytical_solutions_k.png}
\caption{}
\label{fig:numerical_vs_analytical_k}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{figures/numerical_vs_analytical_solutions_a.png}
\caption{}
\label{fig:numerical_vs_analytical_a}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{figures/numerical_vs_analytical_solutions_ak.png}
\caption{}
\label{fig:numerical_vs_analytical_ak}
\end{figure}

>>>>

\begin{equation}
\label{eq:k_short_exact}
\frac{\widetilde{k}}{k} = e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
\end{equation}

@@

\begin{equation}
\label{eq:k_short_lhs}
\frac{\widetilde{k}}{k} = 1 + \varepsilon_L \cos{\psi}
\end{equation}
Coming from the conservation of wave crests, as opposed to the velocity potential
superposition approach, the L-HS solution requires two approximations in their
solution:
First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
second, further truncating the Taylor expansion series beyond the first order term.
These approximations cause an underestimate of the short-wave modulation
magnitude (Fig. \ref{fig:analytical_solutions_ak0.1}a).
At the wave crest, Eq. \ref{eq:k_short_exact} yields the wavenumber modulation
that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of \citet{longuet1960changes}
for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
Although not immediately obvious from the expressions, the modulated wavenumber
in Eq. \ref{eq:k_short_lhs} is conservative across the long-wave phase, whereas
the higher-order solution in Eq. \ref{eq:k_short_exact} is not.
This is due to the fact that combining the orbital velocities evaluated at
$z=\eta$ and the linearized conservation of wave crests in Eq. \ref{eq:wavenumber}
results in excess short-wave wavenumber being created.
The solution of the full Eq. \ref{eq:wavenumber} of course must be conservative
across the long-wave phase, as we will see later from the numerical simulations.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.1.png}
\caption{
  Steady solutions for the modulation of short-wave (a) wavenumber, (b)
  gravitational acceleration, (c) amplitude, (d) steepness, (e) intrinsic
  frequency, and (f) intrinsic phase speed as function of long-wave phase for
  $\varepsilon_L = 0.1$, based on
  \citet{longuet1960changes} (L-HS 1960, black) and this paper (blue).
  Long-wave crest and trough are located at $\psi = 0$ and $\psi = \pi$,
  respectively.
}
\label{fig:analytical_solutions_ak0.1}
\end{figure}

@@

\begin{equation}
\label{eq:gravity_modulation_general}
\frac{\widetilde{g}}{g}
  = \cos{\alpha} 
  + \dfrac{\partial W_{z=\eta}}{\partial t} \cos{\alpha}
  + \dfrac{\partial U_{z=\eta}}{\partial t} \sin{\alpha}
\end{equation}

@@

\begin{equation}
\label{eq:gravity_modulation_linear}
\frac{\widetilde{g}}{g}
  = 
  \frac{
    1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}
    \left[ 1 + \left(\varepsilon_L \sin{\psi}\right)^2 \right]
  }
  {\sqrt{\left(\varepsilon_L \sin{\psi}\right)^2 + 1}}
\end{equation}

@@

\begin{equation}
\label{eq:gravity_modulation_linear_no_curvature}
\frac{\widetilde{g}}{g} = \left( 1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}} \right)
\end{equation}
Further, if we evaluate the orbital accelerations at the mean water level $z=0$,
this further simplifies to:

\begin{equation}
\label{eq:gravity_modulation_linear_no_curvature_mean_level}
\widetilde{g} = \left( 1 - \varepsilon_L \cos{\psi} \right)
\end{equation}
which is the form used by \citet{peureux2021unsteady}, for example.

@@

\begin{equation}
\label{eq:gravity_modulation_stokes}
\frac{\widetilde{g}}{g} =
\left\{
  1 - \varepsilon_L e^{k \eta_{St}}
  \left[ \cos{\psi} -
    \varepsilon_L \sin{\psi} \left(
      \sin{\psi}
      + \varepsilon_L \sin{2\psi}
      - \dfrac{1}{16} \varepsilon_L^2 \sin{\psi}
      + \dfrac{9}{8} \varepsilon_L^2 \sin{3\psi}
    \right)
  \right]
\right\}
\end{equation}
The equivalent expression that includes the curvilinear effects can be derived
by evaluating orbital accelerations and local slope $\alpha$ at $z = \eta_{St}$,
and inserting them into Eq. \ref{eq:gravity_modulation_general}:

\begin{equation}
\label{eq:gravity_modulation_general_stokes}
\frac{\widetilde{g}}{g}
  = \cos{\alpha} 
  + \dfrac{\partial W_{z=\eta_{St}}}{\partial t} \cos{\alpha_{St}}
  + \dfrac{\partial U_{z=\eta_{St}}}{\partial t} \sin{\alpha_{St}}
\end{equation}

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_effective_gravities.png}
\caption{
  (a) Analytical solutions for the effective gravitational acceleration
  modulation by long waves, as function of long-wave phase, for $\varepsilon_L = 0.2$.
  Black is for linear wave evaluated at $z=0$,
  blue and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and without curvature effects,
  and orange and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and with curvature effects.
  (b) As in (a) but normalized by the modulation of the linear wave evaluated at $z=0$, $g_1$.
}
\label{fig:effective_gravities}
\end{figure}

It is instructive to also examine the mean, maximum, and minimum values of the
effective gravity as function of long-wave steepness $\varepsilon_L$
(Fig. \ref{fig:effective_gravities_mean_max_min}).
The long-wave averaged effective gravity of short waves remains equal to the
resting value of $g$ for all values of $\varepsilon_L$, except if the effects
of curvature are considered, in which case the average effective gravity is
reduced.
This reduction, in case of a long Stokes wave, is $\approx 3\%$ for
$\varepsilon_L = 0.2$ and $\approx 14\%$ for $\varepsilon_L = 0.4$.
As the effective gravity minima and maxima occur at crests and troughs,
respectively, the curvature affects the phase-averaged values but not the extremes.
The effective gravity reduction at the crests is $\approx 11\%$, $25\%$, and $66\%$
for $\varepsilon_L = 0.1$, $0.2$, and $0.4$, respectively.
These are similar to the fully-nonlinear numerical estimates of
\citet{longuet1986eulerian,longuet1987propagation}.
Note that these are strictly Eulerian values, evaluated at fixed positions in
space.
As \citet{longuet1986eulerian} pointed out, the effective gravity in the
Lagrangian frame is smoother and less amplified because the moving particles
spend a longer time riding near the crests and less time returning through the
troughs.
However, when we solve the short-wave crest and action conservation equations
in an Eulerian frame, it is the Eulerian effective gravity that must be applied
to conserve the short-wave properties across the long-wave phase.

@@

The modulation of short-wave amplitude can be derived in a similar way as we did
for the wavenumber, except that here we linearize the wave action balance and
integrate wave action in time:

\begin{equation}
\label{eq:wave_action_modulation}
\frac{\widetilde{N}}{N} = e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
\end{equation}

Since wave action is energy divided by the intrinsic frequency $\sigma$,
and energy scales with $ga^2$, the modulation of the amplitude is related to the
modulations of gravity, wavenumber, and action:

\begin{equation}
\label{eq:wave_amplitude_modulation}
\dfrac{\widetilde{a}}{a} = \sqrt{
  \dfrac{g}{\widetilde{g}}
  \dfrac{\widetilde{\sigma}}{\sigma}
  \dfrac{\widetilde{N}}{N}}
=
  \left( \dfrac{g}{\widetilde{g}} \right)^{0.25}
  \left( \dfrac{\widetilde{k}}{k} \right)^{0.25}
  \left( \dfrac{\widetilde{N}}{N} \right)^{0.5}
\end{equation}
To the first order in $\varepsilon_L$, and using Eqs. \ref{eq:k_short_exact}
and \ref{eq:wave_action_modulation}, the above simplifies to:

@@

The amplitude and steepness modulation of the short waves as a function of the
long-wave phase for $\varepsilon_L = 0.1$, $0.2$, and $0.4$ are shown and compared
to the \citet{longuet1960changes} solutions in Figs.
\ref{fig:analytical_solutions_ak0.1}, \ref{fig:analytical_solutions_ak0.2},
and \ref{fig:analytical_solutions_ak0.4}, respectively.
Also shown are the modulation of the short-wave intrinsic frequency and phase
speed in panels (e) and (f), respectively.
As the wavenumber and gravity modulations are out of phase, with the latter
being somewhat larger in magnitude, the overall intrinsic frequency of the
short waves is reduced across the entire long-wave phase, with highest reduction
at the long-wave crests.
For the short-wave phase speed modulation, the wavenumber and gravity modulations
work in tandem and cause the short-wave phase speed to be reduced on the long-wave
crests and increased in the troughs.
We will discuss the implications of this effect later in the context of wind
input into short waves.

@@

To quantify the contribution of the nonlinear terms in Eqs. \ref{eq:wavenumber}
and \ref{eq:wave_action2}, and to evaluate the steadiness assumption of the
analytical solutions, we now proceed to numerically integrate the full set of
crest and action conservation equations.
Equations (\ref{eq:wavenumber})-(\ref{eq:wave_action2}) are discretized using
second-order central finite difference in space and integrated in time using the
forth-order Runge-Kutta method \citep{butcher1996runge}.
The space is divided into 100 grid points.
This is effectively the same equation set and numerical configuration as that of
\citet{peureux2021unsteady}, except that their long-wave orbital velocities are
evaluated at $z=0$ instead of $z=\eta_L$, which neglects the Stokes drift of
short-wave groups.
Their gravity modulation also does not consider the nonlinear or the curvature
effects, however, this difference does not qualitatively affect the results.
Another difference is the choice of the numerical scheme for spatial differences,
which in their case is the MUSCL4 scheme by \citet{kurganov2000new}.
Although the second-order central finite difference is not appropriate for many
numerical problems, in our case it is sufficient because it is conservative and
the fields that we compute derivatives of are smooth and continuous.
We show in the next section that our numerical solutions for infinite long-wave
trains are qualitatively equivalent to those of \citet{peureux2021unsteady}.

Finally, the numerical equations here are integrated in a fixed reference frame,
rather than that moving with the long-wave phase speed.
Either approach produces qualitatively equivalent results, however a fixed
reference frame may allow for a more intuitive interpretation of the results.
This difference is important to keep in mind when comparing the results
of \citet{peureux2021unsteady} and those presented here; in their Fig. 1 the
long-wave phase is fixed and the short waves are moving leftward
with the speed of $C_{pL} - C_g - U$ (neglecting group speed inhomogeneity),
whereas in the figures that follow, the long wave is moving rightward with its
phase speed $C_{pL}$ and the short waves (that is, their energy) are moving
rightward with the speed of $C_g + U$ (neglecting group speed inhomogeneity).

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_k.png}
\caption{}
\label{fig:numerical_vs_analytical_k}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_a.png}
\caption{}
\label{fig:numerical_vs_analytical_a}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_ak.png}
\caption{}
\label{fig:numerical_vs_analytical_ak}
\end{figure}
